\documentclass[11pt]{article}
\usepackage{listings}
\newcommand{\numpy}{{\tt numpy}}    % tt font for numpy

\topmargin -.5in
\textheight 9in
\oddsidemargin -.25in
\evensidemargin -.25in
\textwidth 7in

\begin{document}

% ========== Edit your name here
\author{Francesco Penasa}
\title{Distributed System 1 - synch 3}
\maketitle

\medskip

% ========== Begin answering questions here
\texttt{2020 03 03}

\section{Vector clocks} % (fold)
\label{sec:vector_clocks}
\paragraph{Example: bulletin boards} % (fold)
\label{par:example_bulletin_boards}
Msg and reply should arrive in order.
Just look at the rules to understand if we can accept.
\begin{enumerate}
 	\item condition one
 	\item condition two
 \end{enumerate} 
% paragraph example_bulletin_boards (end)
% section vector_clocks (end)

\section{Global States} % (fold)
\label{sec:global_states}

\subsection{Cut} % (fold)
\label{sub:cut}
The union of the partial histories of all the processes.	
\paragraph{Consistent cut} % (fold)
\label{par:consistent_cut}
send should happen before than receive
if receive happen before send it is \textbf{not} consistent
% paragraph consistent_cut (end)
% subsection cut (end)
\paragraph{Distributed Snapshots} % (fold)
\label{par:distributed_snapshots}
A picture of system how it might have been (In a consistent way).
How to do it:
\begin{enumerate}
	\item Assume FIFO
	\item Any process initiate a snapshot: record your state, send it to all outgoing links, monitor all incoming links.
	\item Upon receiving a token: start recording if you were not recording, stop recording the link in which we have received. 
	\item Recording messages: when all incoming have arrived stop monitoring.
\end{enumerate}
% paragraph distributed_snapshots (end)

\section{TODO} % (fold)
\label{sec:todo}
\textbf{Exercises! 45:} write the list of the nodes table style.
% section todo (end)

% section global_states (end)


\section{Exam questions} % (fold)
\label{sec:exam_questions}
\begin{itemize}
	\item Exercise \textbf{synch 31:} $P_1 8,3,4,5 ;$ $P_2 1,9,4,5 ;$ $P_3 1,3,5,5 ;$ $P_4 1,3,4,9;$
	\item Exercise \textbf{synch 32:} 
		\begin{enumerate}
			\item it's not truly multicast because all 3 starting clocks are different, while in multicast there could be only one different from the others.
			\item $P_2 receive(m4) \neq (6,5,5)$
			\item $P_3 send(m2) \neq (6,4,4)$
			\item links are FIFO but $P_1 receive(m4) < P_1 receive(m2)$ while $P_3 send(m2) > P_3 send(m4)$
		\end{enumerate}
\end{itemize}
% section exam_questions (end)


\end{document}
\grid
\grid