\documentclass[11pt]{article}
\usepackage{listings}
\newcommand{\numpy}{{\tt numpy}}    % tt font for numpy

\topmargin -.5in
\textheight 9in
\oddsidemargin -.25in
\evensidemargin -.25in
\textwidth 7in

\begin{document}

% ========== Edit your name here
\author{Francesco Penasa}
\title{Distributed System 1 - }
\maketitle

\medskip

% ========== Begin answering questions here
\texttt{2020 03 31}

\texttt{slide 27}
\paragraph{Implementing Replication} % (fold)
\label{par:implementing_replication}
How to distribute updates and how to keep replicas consistent?
\paragraph{Epidemic Algorithms} % (fold)
\label{par:epidemic_algorithms}
Based on the theory of infectious disease spreading.
\begin{enumerate}
	\item \textbf{good news}intrinsically distributed and redundant: scalable, fault-tolerant, resilient to topological changes.
	\item \textbf{bad news}probabilistic in nature: cannot provide deterministic guarantee 
\end{enumerate}

\paragraph{Consistency protocols} % (fold)
\label{par:consistency_protocols}
implement consistency model by guaranteeing its properties. Sequential consistency is the model most widely used. 
\begin{enumerate}
	\item Primary-based protocols: remote write (datacenters) where a replica server is in charge to serialize and spread. local-write where the client that actually performs the write becomes the new primary-backup. all based on this idea that one replica is more important than the others
	\item Replicated-write protocols:quorum-based, uses voting to ensure sequential consistency, before a read or write, clients must "get permission" by multiple replicas (quorum) example slide 36.
\end{enumerate}

\section{Middleware} % (fold)
\label{sec:middleware}
a software that allows elements of appllications to inmteroperate across network links, despite differences.
\textbf{middleware provide: hardware independence, interchangeability of key software components, network indepence, operational sabings and administrative savings}.
Remote procedure Call (RPC): enable remote access through the well-known procedure call programming.\\
To pass a parameter value
\begin{enumerate}
	\item serialization (convert in byte to transmit the data)
	\item marshalling (encode with data representation info)
\end{enumerate}
middleware provides automates support.
% section middleware (end)



\end{document}
\grid
\grid