\documentclass[11pt]{article}
\usepackage{listings}
\newcommand{\numpy}{{\tt numpy}}    % tt font for numpy

\topmargin -.5in
\textheight 9in
\oddsidemargin -.25in
\evensidemargin -.25in
\textwidth 7in

\begin{document}

% ========== Edit your name here
\author{Francesco Penasa}
\title{Distributed System 1 - lab 1}
\maketitle

\medskip

% ========== Begin answering 
\texttt{davide.vecchia@unitn.it}
\texttt{timofei.istomin@unitn.it}
We will implement some distributed algorithms seen during the course.
\paragraph{akka} % (fold)
 \label{par:akka}
\texttt{akka.io}
framework for reactive concurrent and distributed applications in Java.
Actors interacts by sending messages to each other, the state change w.r.t. the message.
Messages put in a queue and a dispatcher decide how to handle them.

\paragraph{Actor model} % (fold)
\label{par:actor_model}
\begin{enumerate}
	\item variable
	\item queue
	\item dispatcher
	\item handlers
	\item messages
\end{enumerate}

\paragraph{Akka guarantees} % (fold)
\label{par:akka_guarantees}
\begin{enumerate}
	\item Actor may run in parallel
	\item Message queues are FIFO 
	\item Messages might be lost
	\item No actor is running in more than one thread at a time.
\end{enumerate}

Rules 
\begin{enumerate}
	\item encapsulate! Make sure no object is accessible from multiple actors instances.
	\item Don't send recerences to mutable objects, send copies instead.
	\item Don't use non-final static variable in the actor class nor other classes accessible from the actor
	\item Don't use threads that may access internals of an actor.
	\item 
\end{enumerate}

\paragraph{Akka message} % (fold)
\label{par:akka_message}
Serializable. String is inmutable, we don't need final to be sure that the variable is not tampered.
Serializability: turn an object into a series of bytes.

akka messagges should be: immutable, final and implement serializable


\section{Exercise} % (fold)
\label{sec:exercise}
\subsection{Causal Multicast} % (fold)
\label{sub:causal_multicast}
\begin{enumerate}
	\item We'll create a toy group chat application
	\item There will be a group of actors that send chat messages to the whole group (multicast)
	\item For simplicity: all the actors will run locally 
	\item The chat system should guarentee the property of causal delivery.
\end{enumerate}
% subsection causal_multicast (end)
% section exercise (end)


\end{document}
\grid
\grid