\documentclass[11pt]{article}
\usepackage{listings}
\newcommand{\numpy}{{\tt numpy}}    % tt font for numpy

\topmargin -.5in
\textheight 9in
\oddsidemargin -.25in
\evensidemargin -.25in
\textwidth 7in

\begin{document}

% ========== Edit your name here
\author{Francesco Penasa}
\title{Bioinformatics Resources - Introduction to Biological Databases}
\maketitle

\medskip

% ==========
\texttt{2020 03 12}
Amino acid sequences (DNA/RNA), protein sequences and more complex data stored efficiently in databases to be accessible by who needs them to analyze data and generate new knowledge.

\begin{enumerate}
	\item Primary databases: sequences of nucleotides and aminoacids
	\item Derived and specialized databases (protein domains, structures, genes...)
\end{enumerate}
Information gathered through literature, lab analyses and bioinformatics analyses.
Each database is characterized by a central biological element.
In primary data banks \textbf{each element is uniquely identified} by an accession number.
% paragraph biological_data (end)

\begin{enumerate}
	\item Sequences of nucleotides are represented by \textbf{4 characters} strings.
	\item Sequences of amino acids are represented by \textbf{20 characters} strings.
\end{enumerate}
$ACGT = DNA$ $ACGU = RNA$ amino acids are represented by a triplet of DNA alphabet, mapped into a new character.

\paragraph{Primary datatbases} % (fold)
\label{par:primary_datatbases}
data format differs
\begin{enumerate}
	\item GenBank: just store and arachive only nucleotide sequences. Every GenBank record is identified uniquely by the \textbf{ACCESSION and VERSION} codes.
	\item EMBL datalibrary
	\item DDBJ
\end{enumerate}
% paragraph primary_datatbases (end)

\paragraph{Derived data banks} % (fold)
\label{par:derived_data_banks}
\begin{enumerate}
	\item RefSeq: curated and not redundant collection of DNA, RNA and protein sequences.
\end{enumerate}
% paragraph derived_data_banks (end)

\textbf{All NCBI databases are accessible throught a nuque search engine called EWntrez} \texttt{https://www.ncbi.nlm.nih.gov/search/}

\paragraph{Genome browser} % (fold)
\label{par:genome_browser}
Allow us to browese data at various detail levels.

\subparagraph{Reading the gene structure} % (fold)
\label{subp:reading_the_gene_structure}
\begin{enumerate}
	\item Horizontal line with arrows : \textbf{INTRON}
	\item Dark block: \textbf{EXON}	
	\item Thick and arrowless lines: \textbf{UTR}
\end{enumerate}
% subparagraph reading_the_gene_structure (end)
% paragraph genome_browser (end)

\section{Exercises} % (fold)
\label{sec:exercises}
\begin{enumerate}
	\item Display the RARA gene with the browser. How many alternative transcripts does it have? \textbf{4, the transcript are the lines where RARA is written} 
	\item Considering the first transcript, how many introns and exons?8 9  
	\item Now add to the displayed tracks the GC-percent track (it shows the percentage of GC bases along the sequence). Drag it and move it just under the sequence. \textbf{Mapping and Sequencing GC percent dense}
	\item Now hide the alignment track which is shown by default. \textbf{the alignment track is the one that tells you where we are chr17: we go to Mapping and sequences Base position hide}
	\item Zoom in on the 5th UTR region of the transcript and check if there are any known SNPs in that region. \textbf{zoom in to see the place where the 5th track has a thick line arrowless, to show the SNPs go to variation Common SNPs full, or dbSNP for all the SNP}
\end{enumerate}
\begin{enumerate}
	\item We want to retrieve the 3’UTR sequence for our gene, RARA
	\item be careful about the table refseq curated.
	\item \texttt{http://genome.ucsc.edu/cgi-bin/hgTables}
	\item identifiers paste list -> RARA
	\item output format -> sequence, genomic
	\item change some things and submit
\end{enumerate}
 
\begin{enumerate}
	\item Display the KRAS gene with the browser. How many alternative transcripts does it have ? Considering the first transcript, how many introns and exons? 4; go to region in details; 5; 6
	\item Now add the 1000 Genomes – EUR common variants track to the display. Are there 5'UTR variants in KRAS? \textbf{configure, variation, sequence variance, add}
	\item Now zoom to the second exon of the first isoform of KRAS. How many missense variants are there? just zoom and watch
\end{enumerate}

\begin{enumerate}
	\item Get the Ensembl ID and MGI symbol for all mouse genes on chromosome 19. How many genes are there ?  Mouse genes (GRCm38.p6) > click on Filter, region> chromosome 19, Attributes > Gene, only gene stable id; attributes external MGI symbol; click on count; click results
	\item Now limit this to protein coding genes. How many genes are there ? add filter, gene type, protein coding, count, results
	\item Now save the results to a comma-separated values file. ez.
\end{enumerate}
% section exercises (end)

\end{document}
\grid
\grid